---
title: "Revenue_forecasting using Prophet"
author: "Nachu"
date: "`r Sys.Date()`"
output: html_document
---


This project involves building a revenue forecasting model using R  Prophet package. The model is based on historical revenue data and generates revenue predictions for the year 2025. The forecast is visualized with a clear comparison between historical and forecasted revenue.

Data used was generated with the help of LLM, given the following criteria:
A baseline number of visits that fluctuates around 900 to 1,200 visits per month.
Seasonal adjustments to reflect higher visits in certain months (e.g., flu season in fall/winter) and lower visits in summer (when elective procedures are typically down).
Random noise to introduce variability.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,include=TRUE)
```

```{r}
install.packages("prophet",repos = "https://cran.r-project.org")
install.packages("forecast",repos = "https://cran.r-project.org")
library(prophet)
library(stargazer)
library (ggplot2)
```
```{r}
data <- read.csv("C:\\Users\\Nachu\\Downloads\\RCM_Revenue_Forecast_Synthetic_Data_2022_2024.csv")#read data
```

```{r}
data$Date <- as.Date(data$Date)
data_totrev <- data[, c("Date", "Total_Revenue")]
colnames(data_totrev) <- c("ds", "y")#rename columns reqd by package
model <- prophet(data_totrev)

# Make future predictions for the next 12 months
future <- make_future_dataframe(model, periods = 12, freq = "month")

# Predict future values
forecast <- predict(model, future)

forecast_2025 <- subset(forecast, ds >= as.Date("2025-01-01") & ds <= as.Date("2025-12-31"))

# View the forecasted values
forecast_2025[, "yhat"]

#Present in table
stargazer(
  forecast_2025[, c("ds", "yhat", "yhat_lower", "yhat_upper")],
  covariate.labels = c("Date","Income", "Lower bound","Upper bound"),
  type = "text",
  title = "2025 Revenue Forecast",
  rownames = FALSE,
  summary = FALSE,
  digits = 0
)
```

```{r}
# Combine historical and forecasted data
forecast_plot_data <- rbind(
  data.frame(Date = as.Date(data_totrev$ds), Revenue = data$Total_Revenue, Type = "Actual"),
  data.frame(Date = as.Date(forecast_2025$ds), Revenue = forecast_2025$yhat, Type = "Forecast")
)

#Create new column for x axis labels to be displayed as month and year
forecast_plot_data$Month_Year <- format(forecast_plot_data$Date, "%b %Y") 
forecast_plot_data$Month_Year <- as.factor(forecast_plot_data$Month_Year)
forecast_plot_data$Month_Year <- reorder(forecast_plot_data$Month_Year, forecast_plot_data$Date)



# Create the ggplot
ggplot(forecast_plot_data, aes(x = Month_Year, y = Revenue, color = Type, group = Type)) +
  geom_line(size = 1) +
  scale_color_manual(values = c("Actual" = "darkblue", "Forecast" = "darkred")) +
  scale_x_discrete(breaks = levels(forecast_plot_data$Month_Year)[seq(1, length(levels(forecast_plot_data$Month_Year)), by = 3)])+
  labs(title = "Revenue Forecast (2022-2025)",
       x = "Month",
       y = "Total Revenue",
       color = "Legend") +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1))

```
